# `#30DayMapChallenge`
## Day 2 : Lines

I used R to recreate the map titled **The Ridiculousness Of Conference Tournament Locations, In 6 Maps**  which was originally created by 
Neil Paine and Ella Koeze from [**FiveThirtyEight**](https://fivethirtyeight.com/features/the-ridiculousness-of-conference-tournament-locations-in-6-maps/). 
All of the code used to produce this map is below. 


## Data Query 

First data were obtained from the public NCAA database using bigquery 
```{r setup, eval=FALSE}
library(bigrquery)
library(DBI)
library(tidyverse)
library(ggmap)
library(sf)


# Query NCAA Database -----------------------------------------------------

# Project ID
mbb <- "xxxx-xx-xxxx"

con <- dbConnect(
  bigrquery::bigquery(),
  project = "bigquery-public-data",
  dataset = "ncaa_basketball",
  billing = mbb
)


# Load All Division I Men's Basketball Teams
ncaa_mbb <- dbReadTable(con,"mbb_teams")



# Filter Power 5 (and Big East) Conferences
power6 <- ncaa_mbb %>% filter(conf_alias %in% c("ACC","BIG10","BIG12","BIGEAST","PAC12","SEC"))

# Select Addresses of Power 6 Program Venues for Geocoding ---------------------

# Register Google Geocoder API Key
register_google(key = 'ENTER_YOUR_KEY_HERE')

power6Geo <- power6 %>%
  unite(col = 'addresses', c(venue_address,venue_city,venue_state),sep = ",",remove = FALSE) %>%
  mutate(type = "school")


# loop through addresses to geocode lat/lon
for(i in 1:nrow(power6Geo))
{
  # Print("Working...")
  result <- geocode(power6Geo$addresses[i], output = "latlona")
  power6Geo$lon[i] <- as.numeric(result[1])
  power6Geo$lat[i] <- as.numeric(result[2])
  power6Geo$geoAddress[i] <- as.character(result[3])
}

# Geocode Conference Conference Geographic Centers, Suggested, and Actual Tournament Locations
conf_loc <- read_csv("conf_tourn_locations.csv")

# rename columns to bind with Power6 data
confGeo <- conf_loc %>%
  select(-c(sugg_dist, tourney_dist)) %>%
  gather(alias, city, -conf_alias) %>%
  # create type variable for easier grouping in ggplot
  mutate(
    type = case_when(
      alias == "geog_center" ~ "center",
      alias == "tourney_loc" ~ "tourney",
      alias == "sugg_loc" ~ "suggested"
    )
  )

for(i in 1:nrow(confGeo))
{
  # Print("Working...")
  result <- geocode(confGeo$city[i], output = "latlon")
  confGeo$lon[i] <- as.numeric(result[1])
  confGeo$lat[i] <- as.numeric(result[2])
}

confGeo <- confGeo %>% filter(type != 'suggested')

# Bind Power6 School geographic data with conference geographic data

typ <- list(`1` = "school", `2` = "center", `3` = "tourney")
conf <- list(`1` = "ACC", `2` = "BIG10", `3` = "BIGEAST", `4` = "PAC12", `5` ="SEC",`6` = "BIG12")

power6sf <- power6Geo %>%
  select(conf_alias, type,alias,venue_city,venue_state,lon,lat) %>%
  unite('city',c(venue_city,venue_state),sep = ", ",remove = T) %>%
  bind_rows(confGeo) %>%
  st_as_sf( coords = c('lon','lat'),agr = 'constant',crs = 4326, remove = FALSE) %>%
  mutate(type = recode_factor(type,!!!typ),
         conf_alias = recode_factor(conf_alias, !!!conf)
         ) %>%
  mutate(conf_name = factor(conf_alias,
                            labels = c(
                              "Atlantic Coast Conference:\nBrooklyn",
                              "Big Ten:\nNYC",
                              "Big East:\nNYC",
                              "Pacific 12:\nLas Vegas",
                              "Southeastern Conference:\nSt. Louis",
                              "Big 12:\nKansas City"
                            ))
  )

power6sf <- power6sf %>%
  separate(
    conference,
    into = c('conf_nm', 'tourn_loc'),
    sep = ':\n',
    remove = F
  )  %>%
  mutate(conf_nm = factor(
    conf_alias,
    labels = c(
      "Atlantic Coast Conference",
      "Big Ten",
      "Big East",
      "Pacific 12",
      "Southeastern Conference",
      "Big 12"
    )
  ))


# save prepared data
write_rds(ncaa_mbb,"ncaa_mbb_all_d1.rds")
write_rds(power6, "ncaa_mbb_power6.rds")
write_rds(power6sf,"ncaa_mbb_power6_geocoded.rds")
```


## Map Creation

Next I used `ggplot` ,`sf` , among other packages to produce the map. 

```{r plot, echo=TRUE, message=FALSE, warning=FALSE}

################################################################################
## Project: Richard_NCAA_BB_Map
## Script purpose: Replicate FiveThirtyEight NCAA College Basketball Map
## Date: 26-Sep-2019
## Author:Roland Richard
################################################################################
library(tidyverse)
library(rio)
library(sf)
library(lwgeom)
library(spData)
library(tigris)
library(grid)
library(extrafont)
library(ggrepel)
options(tigris_use_cache = TRUE)

# load Data prepared in "code/Richard_NCAA_BB_Data_Query_Prep.R"
conf_loc <- import("conf_tourn_locations.csv")
ncaa_mbb <- import("ncaa_mbb_all_d1.rds")
power6 <- import("ncaa_mbb_power6.rds")
power6Geo <- import("ncaa_mbb_power6_geocoded.rds")

# Function to create lines using great circles
journeys_to_sf <- function(journeys_data,
                           start_long = start.long,
                           start_lat = start.lat,
                           end_long = end.long,
                           end_lat = end.lat) {
  quo_start_long <- enquo(start_long)
  quo_start_lat <- enquo(start_lat)
  quo_end_long <- enquo(end_long)
  quo_end_lat <- enquo(end_lat)

  journeys_data %>%
    select(
      !! quo_start_long,
      !! quo_start_lat,
      !! quo_end_long,
      !! quo_end_lat
    ) %>%
    transpose() %>%
    map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>%
    map(st_linestring) %>%
    st_sfc(crs = 4326) %>%
    st_sf(geometry = .) %>%
    bind_cols(journeys_data) %>%
    select(everything(), geometry)
}

# Load US State Map -------------------------------------------------------

# US State abbreviations
us <- unique(fips_codes$state)[c(1:51)]
cont <- us[-c(2,12)]

# US State Polygon file
states <- states(cb = TRUE)

# Extract Lower 48 States (No Power 6 Schools in AK or HI)
l48 <- st_as_sf(states) %>%
  filter(STUSPS %in% cont) %>%
  st_transform(crs = 2163)


# Connect school points by conference

conf_schls <- power6Geo %>%
  filter(type == "school") %>%
  select(conf_nm, alias,lon,lat) %>%
  st_drop_geometry() %>%
  set_names(c("conf_nm", "alias", "start.long", "start.lat"))


conf_ctrs <- power6Geo %>%
  filter(type == "center") %>%
  select(conf_nm,lon,lat) %>%
  st_drop_geometry() %>%
  set_names(c("conf_nm", "end.long", "end.lat"))

conf_hubs <-
  left_join(conf_schls, conf_ctrs) %>%
  journeys_to_sf() %>%
  st_segmentize(units::set_units(100, km))


# Create Conference footprint boundary
conf_poly <- power6Geo %>%
  filter(type != "tourney") %>%
  group_by(conf_nm) %>%
  summarise() %>%
  st_cast("POLYGON") %>%
  st_convex_hull() %>%
  st_transform(crs = 2163)

# convert to buffer to surround points
conf_bounds <- conf_poly %>% st_buffer(dist = 35000,endCapStyle = "SQUARE")

# merge with conference location data for labels
tourn_labels <-  power6Geo %>% filter(type == 'tourney') 
tourn_labels <- tourn_labels %>%
  mutate(
    lab_x = case_when(
      conf_alias == "BIG12" ~ -95.1,
      conf_alias == "PAC12" ~ -114.1,
      conf_alias == "SEC" ~ -95.1,
      TRUE ~ -45.1
    ),
    lab_y = case_when(
      conf_alias == "BIG12" ~ 40.1,
      conf_alias == "PAC12" ~ 37.1,
      conf_alias == "SEC" ~ 40.1,
      TRUE ~ 41.0    )
  ) 

# Map Power 6 Programs and Conf. Tournament Locations ---------------------

power6_map <- ggplot() +
  geom_sf(
    data = l48,
    fill = '#303030',
    color = '#E8E8E8',
    lwd = 0.1
  ) +
  geom_sf(
    data = conf_hubs,
    aes(group = conf_nm),
    color = '#FF00FF',
    lwd = 1,
    show.legend = FALSE
  ) +
  geom_sf(
    data = power6Geo,
    aes(
      fill = type,
      color = type,
      size = type
    ),
    shape = 21,
    alpha = 0.9,
    show.legend = "point"
  ) +
  geom_sf(
    data = conf_bounds,
    aes(group = conf_nm),
    fill = '#FF00CC',
    color = '#FFFFFF',
    alpha = 0.20
  ) +
  geom_sf(
    data = tourn_labels,
    fill = "#3300FF",
    color = "#FFFFFF",
    size = 5,
    shape = 21,
    show.legend = "point"
  ) +
 geom_label_repel(
    data  =
      tourn_labels %>% filter(conf_alias %in% c("ACC", "BIG10", "BIGEAST")),
    aes(x = lab_x, y = lab_y, label = tourn_loc,geometry = geometry),
    stat = "sf_coordinates",
    nudge = tourn_labels$lab_x + 15, 
    force = 2, 
    hjust = 0, 
    direction = "y",
    size = 3.9,
    col = "#000000",
    fill = "#FFFFFF",
    fontface = "bold"
  ) +
  geom_label_repel(
    data  =
      tourn_labels %>% filter(conf_alias %in% c("BIG12", "PAC12", "SEC")),
    aes(
      x = lab_x,
      y = lab_y,
      label = tourn_loc,
      geometry = geometry
    ),
    stat = "sf_coordinates",
    size = 3.9,
    force = 5,
    direction    = "both",
    col = "#000000",
    fill = "#FFFFFF",
    fontface = "bold"
  ) +
  scale_fill_manual(
    name = "Location:",
    values =  c("#FF00CC","#FF00CC","#3300FF"),
    labels = c(
      "School",
      "Conference Geographic Center",
      "Conference Tournament Location"
    )
  ) +
  scale_color_manual(
    name = "Location:",
    values =  c("#FF00FF","#FFFFFF","#FFFFFF"),
    labels = c(
      "School",
      "Conference Geographic Center",
      "Conference Tournament Location"
    )
  ) +
  scale_size_manual(
    name = "Location:",
    values = c(2, 6, 5),
    labels = c(
      "School",
      "Conference Geographic Center",
      "Conference Tournament Location"
    )
  ) +
  guides(shape = FALSE,
         colour = guide_legend(override.aes = list(
           shape = 21,
           size = 5,
           fill = c("#FF00CC", "#FF00CC", "#3300FF"),
           color = c("#FF00FF","#FFFFFF","#FFFFFF")
         ))) +
  coord_sf(
    crs = st_crs(2163),
    xlim = c(-2500000, 2500000),
    ylim = c(-2300000, 730000),
    clip = "on"
  ) +
  facet_wrap(.~conf_nm,nrow = 2) +
  labs(title = "Conference championships are far from the conference centers",
       subtitle = "Geographic centers of each of the top six men's college basketball conferences and the\nlocations of each of their 2018 conference tournaments\n",
       caption = " Originally Created by FiveThirtyEight\n Recreated by Roland Richard (@rorich)\n") +
  theme_void(base_family = "Roboto Condensed") +
  theme(
    legend.position = c(0.5, 0),
    legend.justification = c(0, 0),
    legend.direction = "horizontal",
    legend.box.just = "center",
    legend.title = element_text(size = 12, face = "bold", hjust = 1),
    legend.text = element_text(size = 10, hjust = 0.5),
    strip.background = element_blank(),
    strip.text.x = element_text(face = 'bold',  size = 16, hjust = 0.5, vjust = 0.5,margin = margin(0.5,0,0.5,0, "cm")),
    plot.title = element_text(size = 24, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(color = "#454545", size = 18, face = "plain", hjust = 0.5),
    plot.caption = element_text(size = 10, face = "plain", color = "#FFFFFF", hjust = 0, vjust = 1),
    panel.background = element_blank(),
    plot.background = element_rect(fill = "#999999")
  )



```

## Map

```{r map, echo=FALSE, fig.height=9, fig.width=14, message=FALSE, warning=FALSE, paged.print=FALSE}
power6_map
```
